Thesis Title

### 2022-04-08 chapters/chapter_4.tex
update chap4

>>>>

\subsection{Dengue virus reference databases} \label{chap4_sup_database}

We have compiled a database of 3858 complete DENV genomes obtained from the NIAID Virus Pathogen Database and Analysis Resource (ViPR) in October 2019 \citep{pickett_virus_2012} (\url{http://www.viprbrc.org/}).The sequences were distributed unevenly throughout the four DENV serotypes, with DENV-1 being the most represented with 1636 sequences (42.72\%), followed by DENV-2 with 1067 sequences (27.86\%), DENV-3 with 807 sequences (21.07\%), and DENV-4 with 320 sequences (8.36\%). The selection criteria for the search were as follows: a) complete genome sequence only, b) human or mosquito host, c) collection year (1950-2018). Data available from all countries was included and duplicated sequences were removed and only the sequences with sub-type data were kept. A representative of DENV serotype 1 genotype III was introduced (EF457905, recovered from monkey) as no representatives were available with the search criteria used. This genotype is sylvatic and considered extinct \citep{villabona-arenas_worldwide_2013, vasilakis_history_2008}. Additionally, any sample with IUPAC codes in the sequence provided were excluded.

In order to recover the maximum number of DENV reads from the input HTS data in the first mapping step (Figure \ref{fig:chap2_figure1}), we maintained the database with the 3858 complete DENV genomes to retain as much diversity as possible. This database is referred as DENV mapping database and is available on GitHub at \url{https://github.com/B-UMMI/DEN-IM/blob/master/ref/DENV_MAPPING_V3.fasta}.
